---
title: "Getting started with R"
author: "Sandra Godinho Silva"
date: "2021-04"
output: 
  html_document:
    theme: cosmo
    highlight: tango
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 4
  
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# R and RStudio

The term "`R`" is used to refer to both the programming language and the
software that interprets the scripts written using it.

[RStudio](https://rstudio.com) is currently a very popular way to not only write
your R scripts but also to interact with the R software. To function correctly,
RStudio needs R and therefore both need to be installed on your computer.


Please ensure you have completed the installation and setup of RStudio and R.


## Verify correct installation


```{r}
version
```

# The R basics

We can now get started with the R command prompt open.

```{r}
x=2
print(x) ##Print method
class(x)

x=seq(1:10) # Create a vector
class(x)
print(x)
print(x[1]) # First index of vector
print(x[1:5])

y = matrix(nrow=5, ncol=5) # create a 5x5 matrix
print(y)
class(y)
y[1,1] = 5
print(y)
y[,1]= x[1:5]
print(y)
class(y[,1])

y = cbind(seq(1:5), 
seq(1:5),
seq(1:5),
seq(1:5),
seq(1:5))

class(y)


```


## Read table/tab/csv/txt text files:
read.table()
read.csv()
read.delim()

```{r}
cars = read.table('https://raw.githubusercontent.com/sandragodinhosilva/microbiomes2021/main/data/mtcars.csv', header=T, sep = ',') # Read a comma separated values file
head(cars)

cars = read.csv('https://raw.githubusercontent.com/sandragodinhosilva/microbiomes2021/main/data/mtcars.csv')

cars = read.delim('https://raw.githubusercontent.com/sandragodinhosilva/microbiomes2021/main/data/mtcars.csv', sep=',')


```

## Write Files

```{r}
write.table(cars, file='mtcars.tab')
write.csv(cars, file='mtcars.csv')
```

Where are these files now?




# More resources

### More about R

-   [Introduction to R](https://cran.r-project.org/doc/manuals/R-intro.pdf)
-   [R FAQ](https://cran.r-project.org/doc/FAQ/R-FAQ.html)
-   To stay up to date, follow `#rstats` on twitter. Twitter can also be a way
    to get questions answered and learn about useful R packages and tipps (e.g.,
    [@RLangTips])


[Credits](https://datacarpentry.org/R-ecology-lesson/00-before-we-start.html)


### [Home](https://sandragodinhosilva.github.io/microbiomes2021/)









